--- 
title: "Advanced Probabilistic Minecraft Farming"
subtitle: ""
author: ""
date: "`r format(Sys.Date(), format = '%d %b %Y')`"
output:
  bookdown::gitbook:
    config:
      toc:
        collapse: section
        scroll_highlight: true
        before: null
        after: null
      theme: united
  bookdown::pdf_book:
    keep_tex: yes
documentclass: book
link-citations: yes
---

<style type="text/css">
.book .book-body .page-wrapper .page-inner {
  max-width: 90%;
}
</style>

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(AshHaven)
library(dplyr)
library(ggplot2)
library(reshape2)

theme_set(theme_bw())

PURPLE <- "#9970AB"

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```

# Introduction 

This document provides the fully developed mathematical theory underlying the Minecraft farming mechanics. It will develop the principles behind ticks and random ticks and extend those principles to growth models. After establishing the foundations, a discussion of growth mechanics for some farms of interest will be entertained. The mathematical results and principles will be of use in estimating farming production for specific items.

The theory and formulae provided in this document rely heavily on conditional probability, Bayes' Rule, and the Hypergeometric distribution. The conceptual background of the terms listed below may be difficult to decipher when mixed in with the mathematical notation. The reader is encouraged to review _Basic Probabilistic Farming_ to better understand the conceptual background of these terms.

## Definitions

* $S$: The number of growth stages a crop experiences before reaching maturity.
* $\Pr(RTick)$: The probability of a random tick being assigned to a block.
* $TTRT$: (Time to Random Tick) The duration of time until a random tick is assigned to a block measured in minutes.
* $\Pr(Growth | RTick)$: (Growth rate) The probability that a crop will advance a growth stage when assigned a random tick.
* $TTG$: (Time to growth) The duration of time until a crop advances a growth stage measured in minutes.
* $TTM$: (Time to maturity) The duration of time until a crop reaches maturity measured in minutes.
* $ECO$: (Expected Crop Output) The calculated yield of new crop per minute returned by harvesting a mature crop.
* $NSG$: (Net Smelting Gain) The net increase in items smelted after accounting for items of the same type consumed in converting the resource to fuel.
* $PFI$: (Perpetual Furnace Index) The number of blocks of a resource that must be farmed to support continuous and uninterrupted operation of a single furnace.

```{r functions}
# Probability or Random Tick
P_RTICK <- 3 / (16^3)

# Time to Random Tick
TTR_MEAN <- 1 / P_RTICK
```

```{r, child = '01-TickMechanics.Rmd'}
```

```{r, child = '02-GrowthProbability.Rmd'}
```

