# Bamboo Farming

## Farm Metrics

```{r}
stages <- 1
yield <- 1
pr_growth <- growth_probability(4096)
```

```{r, results = 'asis'}
cat(crop_summary(growth_probability = pr_growth, 
                 yield = yield, 
                 stages = stages))
cat("\n")
cat(smelt_summary(total_smelt = 0.25, 
                  self_smelt = 0, 
                  craft_consumption = 1, 
                  yield = yield, 
                  growth_probability = pr_growth, 
                  stages = stages, 
                  label = "Bamboo -> Bamboo"))
```


## Growth Stages

New bamboo growth is harvestable after each growth event. If a bamboo starts at a height of one, when it grows to a height of two, the new block at level to is harvestable. This effectively gives bamboo a single stage of growth. 

Players are not required to harvest at the first stage, however, and may opt to delay harvest until subsequent growth stages are complete. Bamboo will grow to a total height between 12 and 16 blocks tall, meaning it may be treated as having between 11 and 15 growth stages. With no practical method to determine what the maximum height of any individual bamboo plant will be, players would do well to treat bamboo as having 11 growth stages and plan harvests accordingly.

## Optimizing Timed Harvests

Suppose bamboo farm is designed with the intent to perform a timed harvest. That is, the player does not wish to user observer blocks to initiate the harvest, but instead wishes to use a timed device. The player also recognizes that if bamboo is left to grow beyond its eleventh growth cycle, there is potential for it to stop growing and reduce the yield of the farm. In order to reduce loss of yield, it is decided that the harvest action will occur at the time that twenty five percent of the bamboo plants have reached the tenth growth stage--where the bamboo is at a total height of eleven blocks and yields ten crops. 

This design allows for only twenty five percent of the bamboo plants the opportunity to reach a height at which they may stop growing. The time between harvests can be calculated as the time to maturity for bamboo with ten growth stages.

```{r}
ttm_25 <- time_to_maturity(pr_growth, stages = 10, p = 0.25, method = "quantile", unit = "minute")
```

$$TTM_{0.25} = `r ttm_25`$$

The player should set the harvest clock for about `r ttm_25` minutes, or `r ttm_25 %/% 1` minutes and `r round((ttm_25 %% 1) * 60)` seconds.

## Bamboo as Fuel

```{r}
FuelData <- "
Raw    | Crafted | NRaw | Consumed | SmeltCrafted
Bamboo | Bamboo  | 1    | 0        | 0.25
Bamboo | Stick   | 2    | 0        | 0.5
Bamboo | Planks  | 9    | 0        | 3
"

FuelData <- 
    read.csv(text = FuelData, 
             sep = "|") %>% 
    mutate(NSG = mapply(net_smelting_gain, 
                        SmeltCrafted, 
                        Consumed, 
                        NRaw, 
                        SIMPLIFY = TRUE), 
           PFI_Median = mapply(perpetual_furnace_index, 
                               SmeltCrafted, 
                               Consumed, 
                               NRaw, 
                               yield = yield, 
                               growth_probability = pr_growth,
                               stages = stages, 
                               SIMPLIFY = TRUE), 
           PFI_Mean = mapply(perpetual_furnace_index, 
                             SmeltCrafted, 
                             Consumed, 
                             NRaw, 
                             yield = yield, 
                             growth_probability = pr_growth,
                             stages = stages,
                             MoreArgs = list(method = "mean"),
                             SIMPLIFY = TRUE))

FuelData %>% 
    select(-Consumed) %>% 
    mutate(NSG = format(round(NSG, 3), nsmall = 3), 
           PFI_Mean = format(round(PFI_Mean, 3), nsmall = 3),
           PFI_Median = format(round(PFI_Median, 3), nsmall = 1)) %>% 
knitr::kable()
```
