--- 
title: "Probabilistic Minecraft Farming"
subtitle: ""
author: ""
date: "`r format(Sys.Date(), format = '%d %b %Y')`"
output:
  bookdown::gitbook:
    config:
      toc:
        collapse: section
        scroll_highlight: true
        before: null
        after: null
      theme: united
  bookdown::pdf_book:
    keep_tex: yes
documentclass: book
link-citations: yes
---

<style type="text/css">
.book .book-body .page-wrapper .page-inner {
  max-width: 90%;
}
</style>

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(AshHaven)
library(dplyr)
library(ggplot2)

theme_set(theme_bw())

PURPLE <- "#9970AB"

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```

# Introduction 

This document provides the mathematical theory underlying the Minecraft farming mechanics. It will develop the principles behind ticks and random ticks and extend those principles to simple growth models. After establishing the foundations, a discussion of growth mechanics for some farms of interest will be entertained. The mathematical results and principles will be of use in estimating farming production for specific items.

## Definitions

* $S$: The number of growth stages a crop experiences before reaching maturity.
* $\Pr(RTick)$: The probability of a random tick being assigned to a block.
* $TTRT$: (Time to Random Tick) The duration of time until a random tick is assigned to a block measured in minutes.
* $\Pr(Growth | RTick)$: (Growth rate) The probability that a crop will advance a growth stage when assigned a random tick.
* $TTG$: (Time to growth) The duration of time until a crop advances a growth stage measured in minutes.
* $TTM$: (Time to maturity) The duration of time until a crop reaches maturity measured in minutes.
* $ECO$: (Expected Crop Output) The calculated yield of new crop per minute returned by harvesting a mature crop.
* $NSG$: (Net Smelting Gain) The net increase in items smelted after accounting for items of the same type consumed in converting the resource to fuel.
* $PFI$: (Perpetual Furnace Index) The number of blocks of a resource that must be farmed to support continuous and uninterrupted operation of a single furnace.

```{r functions}
# Probability or Random Tick
P_RTICK <- 3 / (16^3)

# Time to Random Tick
TTR_MEAN <- 1 / P_RTICK
TTR_MEDIAN <- -1 / (log2(1 - P_RTICK))
TTR_Q <- function(p){ qgeom(p, prob = P_RTICK)}

# Time to Growth
TTG_MEAN <- function(pr_growth, ticks = 1200){ (TTR_MEAN / (pr_growth)) / ticks }
TTG_MEDIAN <- function(pr_growth, ticks = 1200){ (-1 * TTR_MEDIAN / (log2(1 - pr_growth))) / ticks}
TTG_Q <- function(p, pr_growth, ticks = 1200){ (TTR_Q(p) * qgeom(p, prob = pr_growth)) / ticks }

# Time to Maturity
TTM_MEAN <- function(pr_growth, stage, ticks = 1200){TTG_MEAN(pr_growth, ticks) * stage}
TTM_MEDIAN <- function(pr_growth, stage, ticks = 1200){TTG_MEDIAN(pr_growth, ticks) * stage}
TTM_Q <- function(p, pr_growth, stage, ticks = 1200){TTG_Q(p, pr_growth, ticks) * stage}

# Expected Crop Output
ECO_MEAN <- function(yield, pr_growth, stage, ticks = 1200){
  yield / TTM_MEAN(pr_growth, stage, ticks)
}

ECO_MEDIAN <- function(yield, pr_growth, stage, ticks = 1200){
  yield / TTM_MEDIAN(pr_growth, stage, ticks)
}

ECO_Q <- function(yield, pr_growth, stage, ticks = 1200){
  yield / TTM_Q(pr_growth, stage, ticks)
}

# Net Smelting Gain 
NSG <- function(OI, SI, CC){
  (OI - SI) / CC
}

# Perpetual Furnace Index
PFI <- function(nsg, eco){
  (6) / (nsg * eco)
}
```

```{r, child = '00-SupportFunction.Rmd'}
```

```{r, child = '01-TickMechanics.Rmd'}
```

```{r, child = '02-GrowthProbability.Rmd'}
```

```{r, child = '03-TimeToGrowth.Rmd'}
```

```{r, child = '04-TimeToMaturity.Rmd'}
```

```{r, child = '05-ExpectedCropOutput.Rmd'}
```

```{r, child = '06-NetSmeltingGain.Rmd'}
```

```{r, child = '07-PerpetualFurnaceIndex.Rmd'}
```

```{r, child = 'BambooFarming.Rmd'}
```

```{r, child = 'KelpFarming.Rmd'}
```

```{r, child = 'WheatFarming.Rmd'}
```

```{r, child = 'WheatSeedFarming.Rmd'}
```
